Hierarchy Builder: algebraic~hierarchies made easy in Coq with Elpi

### 2020-05-04 hb.tex
Consistent variable names

The first version of our hierarchy (that we name V1) features only two
structures: \coq{Monoid} and \coq{Ring}. We use notations from Appendix~\ref{sec:coq-reference} to define them.
%\marginpar{remove hb.end for \mixins{}, use next release of coq elpi}
\begin{coqcode}
#\hbmixin{}# Record Monoid_of_Type A := {                       #\label{demo:mixin:mot}#
  zero : A;
  add : A -> A -> A;
  addrA : associative add;             (* `add` is associative.                   *)
  add0r : left_id zero add;            (* `zero` is the left and right neutral    *)
  addr0 : right_id zero add;           (*   element with respect to `add`.        *)
}.                                                             #\label{demo:mixin:mot:end}#
#\hbstructure{}# Definition Monoid := { A of Monoid_of_Type A }.        #\label{demo:structure:monoid}#

#\hbmixin{}# Record Ring_of_Monoid A of Monoid A := {       #\label{demo:mixin:rom}#
  one : A;
  opp : A -> A;
  mul : A -> A -> A;
  addNr : left_inverse zero opp add;   (* `opp x` is the left and right additive  *)
  addrN : right_inverse zero opp add;  (*   inverse of `x`.                       *)
  mulrA : associative mul;             (* `mul` is associative.                   *)
  mul1r : left_id one mul;             (* `one` is the left and right neutral     *)
  mulr1 : right_id one mul;            (*   element with respect to `mul`.        *)
  mulrDl : left_distributive mul add;  (* `mul` is left and right distributive    *)
  mulrDr : right_distributive mul add; (*   over `add`.                           *)
}.
#\hbstructure{}# Definition Ring := { A of Monoid A & Ring_of_Monoid A }. #\label{demo:structure:ring}#

@@

\Mixins{} are declared via the \hbmixin{} command
that takes a record declaration with
a type parameter and a possibly empty list of \factories{} for that type.
The code between lines~\ref{demo:mixin:mot}
and~\ref{demo:mixin:mot:end} declares a \mixin{} that can turn a naked
type \coq{A} into a monoid, hence we chose the name to be \coq{Monoid_of_Type}.

The \hbstructure{} command takes in input a definition for a sigma type S
that equips \coq{A} with a list of \factories{}.
It registers a structure S in the hierarchy placing
any definition specific to that structure inside a Coq module named \coq{S}.
Line~\ref{demo:structure:monoid} hence forges the \coq{Monoid} structure.

Line~\ref{demo:mixin:rom} declares a second \mixin{} collecting the operations
and properties that are needed in order to enrich a monoid to a ring, hence
the name \coq{Ring_of_Monoid}. Indeed this time the type variable \coq{A}
is followed by \coq{Monoid} that enriches
\coq{A} with the operations and properties of monoids. As a consequence
\coq{add} and \coq{zero} can be used to express the new properties.

@@

We can print the type for monoids as forged by \HB{} (line~\ref{demo:theory:print:type}).
It packs a carrier, called \coq{sort}, and the collection of operation and
properties that we omit for brevity.  We can also look at the type of two
constants synthesized by \HB{} out of the hierarchy declaration. Remark that
while the names of the constants come from the names of \mixin{} fields,
their types differ.
In particular they are quantified over a \coq{Monoid.type} or \coq{Ring.type},
and not a simple type \coq{A} as in the \mixins{}. Moreover we evince that
the \coq{sort} projection is declared as an implicit coercion~\cite{Saibi97}
and is automatically inserted in order to make \coq{M -> M -> M} a meaningful
type for binary operations on the carrier of \coq{M}. Last, we see that
properties are quantified on (hence apply to) the structure they belong to but
use, in their statements, operations belonging to simpler structures.
For example \coq{addNr} is a property of a ring but  its statement mentions \coq{add},\
the operation of the underlying monoid.

@@

\begin{coqcode}
#\hbmixin{}# Record Monoid_of_Type A := { ... (* unchanged *) ... }.
#\hbstructure{}# Definition Monoid := { A of Monoid_of_Type A }.

@@

#\hbmixin{}# Record Ring_of_AbelianGroup A of AbelianGroup A := {
  one : A;
  mul : A -> A -> A;
  mulrA : associative mul;
  mul1r : left_id one mul;              mulr1 : right_id one mul;
  mulrDl : left_distributive mul add;   mulrDr : right_distributive mul add;
}.
#\hbstructure{}# Definition Ring := { A of AbelianGroup A & Ring_of_AbelianGroup A }.

Lemma addrN {R : AbelianGroup.type} : right_inverse zero opp add.               #\label{demo2:proof:addrN}#
Proof. by move=>x; rewrite addrC addNr. Qed.
\end{coqcode}

Some operations and properties were moved from the old \mixin{} for rings
into a newborn \mixin{} \coq{AbelianGroup_of_Monoid} that gathers the
axioms needed to turn a monoid into an Abelian group. Consequently the \mixin{}
for rings is now called \coq{Ring_of_AbelianGroup} (instead of \coq{Ring_of_Monoid})
since it expects the type \coq{A} to be already an Abelian group and hence
gathers fewer axioms.
While operations moved from one structure to another, some properties
undergo a deep change in their status.
The lemma \coq{addrC} part of the abstract theory of
rings is now an axiom of Abelian groups, while \coq{addrN} is no more
an axiom of rings, but rather a theorem of the abstract theory of Abelian
groups proved at line~\ref{demo2:proof:addrN}.

@@

\begin{coqcode}
#\hbfactory{}# Record Ring_of_Monoid A of Monoid A := { ... (* unchanged *) ...}.

#\hbbuilders{}# Context A (f : Ring_of_Monoid A).                          #\label{demo3:variable:f}#

@@

  Definition to_AbelianGroup_of_Monoid :=
    AbelianGroup_of_Monoid.Build A (opp f) addrC (addNr f). (* addrN unused *)
  #\hbinstance{}# A to_AbelianGroup_of_Monoid.

  Definition to_Ring_of_AbelianGroup :=  Ring_of_AbelianGroup.Build A
    (one f) (mul f) (mulrA f) (mul1r f) (mulr1 f) (mulrDl f) (mulrDr f).
  #\hbinstance{}# A to_Ring_of_AbelianGroup.

@@

The record \coq{Ring_of_Monoid} is the same we declared as a \mixin{}
in version V1. In order to make a \factory{} out of it we equip it with
two definitions that embody the \emph{\builder{}s}.
The first is \coq{to_AbelianGroup_of_Monoid} and
explains how to build an \coq{AbelianGroup} structure out of the \factory{}
axioms (named \coq{f}, line~\ref{demo3:variable:f}).
This construction is also registered as canonical for \coq{A},
so that the next construction \coq{to_Ring_of_AbelianGroup} can call
the \coq{Ring_of_AbelianGroup.Build} \phantterm{} that requires \coq{A} to be
an \coq{AbelianGroup}.
It is worth pointing out that the proof of \coq{addrC} we had in V1 is now
required in order to write the \builder{} for Abelian groups, while
the \coq{addrN} field is not used (the same statement is already part of
the theory of Abelian groups, see line~\ref{demo2:proof:addrN} of the previous
code snippet).

@@

\begin{coqcode}
#\hbstructure{}# Definition Ring := { A of Monoid A & Ring_of_Monoid A }.
#\hbstructure{}# Definition Ring := { A of AbelianGroup A & Ring_of_AbelianGroup A }.
\end{coqcode}

@@

We describe mathematical structures with three kinds of dependent records: mixins, classes, and structures.
As shown in Section \ref{sec:example}, a \mixin{} gathers operators and axioms newly introduced by a structure.
As in \cite[Section 2.4]{DBLP:conf/tphol/GarillotGMR09}\cite[Section 2]{KSdraft}, a class record is parametrized by the carrier type \coq{(T : Type)} and gathers all the operators and axioms of a structure by assembling mixins, and a \coq{Structure type} (a record) bundles a carrier and its class instance, as follows.
\begin{coqcode}
Module Monoid.
Record axioms (T : Type) : Type :=
  Class { Monoid_of_Type_mixin : Monoid_of_Type T; }.
Structure type : Type := Pack { sort : Type; class : axioms sort; }.
End Monoid.
\end{coqcode}
The \coq{Monoid} module plays the role of a name space and forces us to write
qualified names such as \coq{Monoid.type}; as a consequence we can reuse the
same unqualified names for other structures, i.e., class and structure record
can always be named as \coq{axioms} and \coq{type} respectively.

@@

Definition zero {T : Monoid.type} : T :=
  Monoid_of_Type.zero T (Monoid.Monoid_of_Type_mixin T (Monoid.class T)).
Definition add {T : Monoid.type} : T -> T -> T :=
  Monoid_of_Type.add T (Monoid.Monoid_of_Type_mixin T (Monoid.class T)).
Lemma addrA {T : Monoid.type} : associative (@add T).
(* Two monoid axioms `add0r` and `addr0` are omitted. *)
\end{coqcode}

Next we define Abelian groups.
Since the monoid structure is the bottom of this hierarchy its class record
\coq{Monoid} consists of just one \mixin{}. In contrast the class record
of Abelian groups consists of two \mixins{} where the second one depends
 on the former (since Abelian groups inherit from monoids).
\begin{coqcode}
Module AbelianGroup.
Record axioms (T : Type) : Type := Class {
  Monoid_of_Type_mixin : Monoid_of_Type T;
  AbelianGroup_of_Monoid_mixin : AbelianGroup_of_Monoid T Monoid_of_Type_mixin; }.
Structure type : Type := Pack { sort : Type; class : axioms sort }.
End AbelianGroup.
\end{coqcode}
In the flat variant of Packed Classes, a class record gathers mixins as its
fields directly as in above. In contrast, a class record of the non-flat variant
packs a class instance of one of the superclasses with remaining mixins, which
reduces the amount of code to implement inheritance significantly.
%Nested class records and chains of their projections exponentially slow down the
%conversion while the flat variant can mitigate the issue. 
Since we do not need
to care about the amount of code in automated generation, \hb{} uses the flat
variant of Packed Classes as its target language.

As in the monoid structure, we declare \coq{AbelianGroup.sort} as an implicit
coercion and lift additive inverse \coq{opp} and Abelian group axioms.
Since Abelian groups inherit from monoids, we also declare an implicit coercion
from Abelian groups to monoids. A coercion between structures can be defined
in two steps: first a coercion between the class record of the
superclass to the class record of the subclass (line~\ref{demo4:coe:ag:monoid:class});
and a coercion between
structure records (line~\ref{demo4:coe:ag:monoid:type}) relying on the first one.
\begin{coqcode}
Coercion AbelianGroup.sort : AbelianGroup.type >-> Sortclass.
Coercion AbelianGroup_class_to_Monoid_class (T : Type) (c : AbelianGroup T) : #\label{demo4:coe:ag:monoid:class}#
  Monoid T := Monoid.Class T (AbelianGroup.Monoid_of_Type_mixin T c).
Coercion AbelianGroup_to_Monoid (T : AbelianGroup.type) : Monoid.type := #\label{demo4:coe:ag:monoid:type}#
  Monoid.Pack (AbelianGroup.sort T) (AbelianGroup.class T).

Definition opp {T : AbelianGroup.type} : T -> T := ... .
(* Two Abelian group axioms `addrC` and `addNr` are omitted. *)
\end{coqcode}

Generally, a coercion from a structure A to another structure B can (and should) have the following form, thanks to the corresponding coercion between classes \coq{A_class_to_B_class}.
\begin{coqcode}
Coercion A_to_B (T : A.type) : B.type :=
  B.Pack (A.sort T) ((* A_class_to_B_class _ *) (A.class T)).
\end{coqcode}

@@

\begin{coqcode}
#\hbmixin{}# Record SemiRing_of_Monoid A
  of Monoid A := {
  one : A;
  mul : A -> A -> A;
  (* 7 axioms are omitted. *)
}.

#\hbfactory{}# Record Ring_of_AbelianGroup A
  of AbelianGroup A := {
  (* 2 operators and 5 axioms are omitted. *)
}.
#\hbbuilders{}# Context A (a : Ring_of_AbelianGroup A).
..
#\hbend{}#.
\end{coqcode}

Since semirings and Abelian groups do not inherit from each other,
the definition of semirings and Abelian groups are quite similar:
\begin{coqcode}
Module SemiRing.
Record axioms (T : Type) : Type := Class {
  Monoid_of_Type_mixin : Monoid_of_Type T;
  SemiRing_of_Monoid_mixin : SemiRing_of_Monoid T Monoid_of_Type_mixin;
}.
Structure type : Type := Pack { sort : Type; class : axioms sort; }.
End SemiRing.
\end{coqcode}
We define implicit coercions from the semiring structure to the carrier
and the monoid structure, and we lift semiring operators and axioms as follows:
\begin{coqcode}
Coercion SemiRing.sort : SemiRing.type >-> Sortclass.
Coercion SemiRing_to_Monoid : SemiRing.type >-> Monoid.type.

Definition one {T : SemiRing.type} : T := ... .
Definition mul {T : SemiRing.type} : T -> T -> T := ... .
(* 7 semiring axioms are omitted. *)
\end{coqcode}

The class record is defined by gathering the monoid, Abelian group, and
semiring \mixins{}. Since the rings inherit from monoids, semirings, and
Abelian groups, we define implicit coercions from the ring structure to those
three structures.
\begin{coqcode}
Module Ring.
Record axioms (T : Type) : Type := Class {
  Monoid_of_Type_mixin : Monoid_of_Type T;
  SemiRing_of_Monoid_mixin : SemiRing_of_Monoid T Monoid_of_Type_mixin;
  AbelianGroup_of_Monoid_mixin : AbelianGroup_of_Monoid T Monoid_of_Type_mixin; }.
Structure type : Type := Pack { sort : Type; class : axioms sort; }.
End Ring.

@@

Another example of elaboration problem is \(- 1\), which hides the term~\coq{opp _ (one _)}.
Here \coq{opp} and \coq{one} are respectively from Abelian groups and semirings,
which do not inherit each other but whose smallest common substructure is
rings; thus we have to extend the unifier to solve a unification problem
\(\coq{AbelianGroup.sort}~?_{\texttt{AbG}}~\coq{= SemiRing.sort}~?_{\texttt{SR}}\) by instantiating \(?_{\texttt{AbG}}\) and \(?_{\texttt{SR}}\) with \(\coq{Ring_to_AbelianGroup}~?_{\texttt{R}}\) and \(\coq{Ring_to_SemiRing}~?_{\texttt{R}}\) respectively where \(?_{\texttt{R}}\) is a fresh unification variable of type \coq{Ring.type}.
This hint can be given as follows:% can be done by the following canonical declaration.
\begin{coqcode}
Canonical AbelianGroup_to_SemiRing (T : Ring.type) :=
  SemiRing.Pack (AbelianGroup.sort (Ring_to_AbelianGroup T)) (Ring.class T)).
\end{coqcode}

@@

(* Bottom mixin in Fig. 2. *)
HB.mixin Record Monoid_of_Type A := {
  zero : A;
  add : A -> A -> A;
  addrA : associative add;
  add0r : left_id zero add;
  addr0 : right_id zero add;
}.
HB.structure Definition Monoid := { A of Monoid_of_Type A }.
Notation "0" := zero : hb_scope.
Infix "+" := (@add _) : hb_scope.

@@

(* Bottom left mixin in Fig. 2. *)
HB.mixin Record SemiRing_of_Monoid A of Monoid A := {
  one : A;
  mul : A -> A -> A;
  mulrA : associative mul;
  mul1r : left_id one mul;
  mulr1 : right_id one mul;
  mulrDl : left_distributive mul add;
  mulrDr : right_distributive mul add;
  mul0r : left_zero zero mul;
  mulr0 : right_zero zero mul;
}.
HB.structure Definition SemiRing := { A of Monoid A & SemiRing_of_Monoid A }.
Notation "1" := one : hb_scope.
Infix "*" := (@mul _) : hb_scope.

Lemma addrN {R : AbelianGroup.type} : right_inverse (zero : R) opp add.
Proof. by move=> x; rewrite addrC addNr. Qed.

(* Top right factory in Fig. 2. *)
HB.factory Record Ring_of_AbelianGroup A of AbelianGroup A := {
  one : A;
  mul : A -> A -> A;
  mulrA : associative mul;
  mul1r : left_id one mul;
  mulr1 : right_id one mul;
  mulrDl : left_distributive mul add;
  mulrDr : right_distributive mul add;
}.

@@

(* Top left factory in Fig. 2. *)
(* It is an exact copy of the bottom right mixin. *)
HB.factory Definition Ring_of_SemiRing A of SemiRing A := AbelianGroup_of_Monoid A.
(* The corresponding builder is the identity. *)
HB.builders Context (A : Type) (f : Ring_of_SemiRing A).
  Definition to_AbelianGroup_of_Monoid : AbelianGroup_of_Monoid A := f.
  HB.instance A to_AbelianGroup_of_Monoid.
HB.end.

(* Right-most factory in Fig. 2. *)
HB.factory Record Ring_of_Monoid A of Monoid A := {
  one : A;
  opp : A -> A;
  mul : A -> A -> A;
  addNr : left_inverse zero opp add;
  addrN : right_inverse zero opp add;
  mulrA : associative mul;
  mul1r : left_id one mul;
  mulr1 : right_id one mul;
  mulrDl : left_distributive mul add;
  mulrDr : right_distributive mul add;
}.

>>>>

The first version of our hierarchy (that we name V1) features only two
structures: \coq{Monoid} and \coq{Ring}. We use notations from Appendix~\ref{sec:coq-reference} to define them.
%\marginpar{remove hb.end for \mixins{}, use next release of coq elpi}
\begin{coqcode}
#\hbmixin{}# Record Monoid_of_Type M := {                       #\label{demo:mixin:mot}#
  zero : M;
  add : M -> M -> M;
  addrA : associative add;             (* `add` is associative.                   *)
  add0r : left_id zero add;            (* `zero` is the left and right neutral    *)
  addr0 : right_id zero add;           (*   element with respect to `add`.        *)
}.                                                             #\label{demo:mixin:mot:end}#
#\hbstructure{}# Definition Monoid := { M of Monoid_of_Type M }.        #\label{demo:structure:monoid}#

#\hbmixin{}# Record Ring_of_Monoid R of Monoid R := {       #\label{demo:mixin:rom}#
  one : R;
  opp : R -> R;
  mul : R -> R -> R;
  addNr : left_inverse zero opp add;   (* `opp x` is the left and right additive  *)
  addrN : right_inverse zero opp add;  (*   inverse of `x`.                       *)
  mulrA : associative mul;             (* `mul` is associative.                   *)
  mul1r : left_id one mul;             (* `one` is the left and right neutral     *)
  mulr1 : right_id one mul;            (*   element with respect to `mul`.        *)
  mulrDl : left_distributive mul add;  (* `mul` is left and right distributive    *)
  mulrDr : right_distributive mul add; (*   over `add`.                           *)
}.
#\hbstructure{}# Definition Ring := { R of Monoid R & Ring_of_Monoid R }. #\label{demo:structure:ring}#

@@

\Mixins{} are declared via the \hbmixin{} command
that takes a record declaration with
a type parameter and a possibly empty list of \factories{} for that type.
The code between lines~\ref{demo:mixin:mot}
and~\ref{demo:mixin:mot:end} declares a \mixin{} that can turn a naked
type \coq{M} into a monoid, hence we chose the name to be \coq{Monoid_of_Type}.

The \hbstructure{} command takes in input a definition for a sigma type \coq{S}
that equips a type with a list of \factories{}.
It registers a structure \coq{S} in the hierarchy placing
any definition specific to that structure inside a Coq module named \coq{S}.
Line~\ref{demo:structure:monoid} hence forges the \coq{Monoid} structure.

Line~\ref{demo:mixin:rom} declares a second \mixin{} collecting the operations
and properties that are needed in order to enrich a monoid to a ring, hence
the name \coq{Ring_of_Monoid}. Indeed this time the type variable \coq{R}
is followed by \coq{Monoid} that enriches
\coq{R} with the operations and properties of monoids. As a consequence
\coq{add} and \coq{zero} can be used to express the new properties.

@@

We can print the type for monoids as forged by \HB{} (line~\ref{demo:theory:print:type}).
It packs a carrier, called \coq{sort}, and the collection of operation and
properties that we omit for brevity.  We can also look at the type of two
constants synthesized by \HB{} out of the hierarchy declaration. Remark that
while the names of the constants come from the names of \mixin{} fields,
their types differ.
In particular they are quantified over a \coq{Monoid.type} or \coq{Ring.type},
and not a simple type as in the \mixins{}. Moreover we evince that
the \coq{sort} projection is declared as an implicit coercion~\cite{Saibi97}
and is automatically inserted in order to make \coq{M -> M -> M} a meaningful
type for binary operations on the carrier of \coq{M}. Last, we see that
properties are quantified on (hence apply to) the structure they belong to but
use, in their statements, operations belonging to simpler structures.
For example \coq{addNr} is a property of a ring but  its statement mentions \coq{add},\
the operation of the underlying monoid.

@@

\begin{coqcode}
#\hbmixin{}# Record Monoid_of_Type M := { ... (* unchanged *) ... }.
#\hbstructure{}# Definition Monoid := { M of Monoid_of_Type M }.

@@

#\hbmixin{}# Record Ring_of_AbelianGroup R of AbelianGroup R := {
  one : R;
  mul : R -> R -> R;
  mulrA : associative mul;
  mul1r : left_id one mul;              mulr1 : right_id one mul;
  mulrDl : left_distributive mul add;   mulrDr : right_distributive mul add;
}.
#\hbstructure{}# Definition Ring := { R of AbelianGroup R & Ring_of_AbelianGroup R }.

Lemma addrN {A : AbelianGroup.type} : right_inverse zero opp add.               #\label{demo2:proof:addrN}#
Proof. by move=> x; rewrite addrC addNr. Qed.
\end{coqcode}

Some operations and properties were moved from the old \mixin{} for rings
into a newborn \mixin{} \coq{AbelianGroup_of_Monoid} that gathers the
axioms needed to turn a monoid into an Abelian group. Consequently the \mixin{}
for rings is now called \coq{Ring_of_AbelianGroup} (instead of \coq{Ring_of_Monoid})
since it expects the type \coq{R} to be already an Abelian group and hence
gathers fewer axioms.
While operations moved from one structure to another, some properties
undergo a deep change in their status.
The lemma \coq{addrC} part of the abstract theory of
rings is now an axiom of Abelian groups, while \coq{addrN} is no more
an axiom of rings, but rather a theorem of the abstract theory of Abelian
groups proved at line~\ref{demo2:proof:addrN}.

@@

\begin{coqcode}
#\hbfactory{}# Record Ring_of_Monoid R of Monoid R := { ... (* unchanged *) ...}.

#\hbbuilders{}# Context R (f : Ring_of_Monoid R).                          #\label{demo3:variable:f}#

@@

  Definition to_AbelianGroup_of_Monoid :=
    AbelianGroup_of_Monoid.Build R opp addrC addNr. (* addrN unused *)
  #\hbinstance{}# R to_AbelianGroup_of_Monoid.

  Definition to_Ring_of_AbelianGroup := Ring_of_AbelianGroup.Build R one mul
    mulrA mul1r mulr1 mulrDl mulrDr.
  #\hbinstance{}# R to_Ring_of_AbelianGroup.

@@

The record \coq{Ring_of_Monoid} is the same we declared as a \mixin{}
in version V1. In order to make a \factory{} out of it we equip it with
two definitions that embody the \emph{\builder{}s}.
The first is \coq{to_AbelianGroup_of_Monoid} and
explains how to build an \coq{AbelianGroup} structure out of the \factory{}
axioms (named \coq{f}, line~\ref{demo3:variable:f}).
This construction is also registered as canonical for \coq{R},
so that the next construction \coq{to_Ring_of_AbelianGroup} can call
the \coq{Ring_of_AbelianGroup.Build} \phantterm{} that requires \coq{R} to be
an \coq{AbelianGroup}.
It is worth pointing out that the proof of \coq{addrC} we had in V1 is now
required in order to write the \builder{} for Abelian groups, while
the \coq{addrN} field is not used (the same statement is already part of
the theory of Abelian groups, see line~\ref{demo2:proof:addrN} of the previous
code snippet).

@@

\begin{coqcode}
#\hbstructure{}# Definition Ring := { R of Monoid R & Ring_of_Monoid R }.
#\hbstructure{}# Definition Ring := { R of AbelianGroup R & Ring_of_AbelianGroup R }.
\end{coqcode}

@@

We describe mathematical structures with three kinds of dependent records: mixins, classes, and structures.
As shown in Section \ref{sec:example}, a \mixin{} gathers operators and axioms newly introduced by a structure.
As in \cite[Section 2.4]{DBLP:conf/tphol/GarillotGMR09}\cite[Section 2]{KSdraft}, a class record is parametrized by the carrier type \coq{(T : Type)} and gathers all the operators and axioms of a structure by assembling mixins, and a \coq{Structure type} (a record) bundles a carrier and its class instance, as follows.
\begin{coqcode}
Module Monoid.
Record axioms (M : Type) : Type :=
  Class { Monoid_of_Type_mixin : Monoid_of_Type M; }.
Structure type : Type := Pack { sort : Type; class : axioms sort; }.
End Monoid.
\end{coqcode}
The \coq{Monoid} module plays the role of a name space and forces us to write
qualified names such as \coq{Monoid.type}; as a consequence we can reuse the
same unqualified names for other structures, i.e., class and structure record
can always be named as \coq{axioms} and \coq{type} respectively.

@@

Definition zero {M : Monoid.type} : M :=
  Monoid_of_Type.zero M (Monoid.Monoid_of_Type_mixin M (Monoid.class M)).
Definition add {M : Monoid.type} : M -> M -> M :=
  Monoid_of_Type.add M (Monoid.Monoid_of_Type_mixin M (Monoid.class M)).
Lemma addrA {M : Monoid.type} : associative (@add M).
(* Two monoid axioms `add0r` and `addr0` are omitted. *)
\end{coqcode}

Next we define Abelian groups.
Since the monoid structure is the bottom of this hierarchy its class record
\coq{Monoid} consists of just one \mixin{}. In contrast the class record
of Abelian groups consists of two \mixins{} where the second one depends
 on the former (since Abelian groups inherit from monoids).
\begin{coqcode}
Module AbelianGroup.
Record axioms (A : Type) : Type := Class {
  Monoid_of_Type_mixin : Monoid_of_Type A;
  AbelianGroup_of_Monoid_mixin : AbelianGroup_of_Monoid A Monoid_of_Type_mixin; }.
Structure type : Type := Pack { sort : Type; class : axioms sort }.
End AbelianGroup.
\end{coqcode}
In the flat variant of Packed Classes, a class record gathers mixins as its
fields directly as in above. In contrast, a class record of the non-flat variant
packs a class instance of one of the superclasses with remaining mixins, which
reduces the amount of code to implement inheritance significantly.
%Nested class records and chains of their projections exponentially slow down the
%conversion while the flat variant can mitigate the issue. 
Since we do not need
to care about the amount of code in automated generation, \hb{} uses the flat
variant of Packed Classes as its target language.

As in the monoid structure, we declare \coq{AbelianGroup.sort} as an implicit
coercion and lift additive inverse \coq{opp} and Abelian group axioms.
Since Abelian groups inherit from monoids, we also declare an implicit coercion
from Abelian groups to monoids. A coercion between structures can be defined
in two steps: first a coercion between the class record of the
superclass to the class record of the subclass (line~\ref{demo4:coe:ag:monoid:class});
and a coercion between
structure records (line~\ref{demo4:coe:ag:monoid:type}) relying on the first one.
\begin{coqcode}
Coercion AbelianGroup.sort : AbelianGroup.type >-> Sortclass.
Coercion AbelianGroup_class_to_Monoid_class (A : Type) (c : AbelianGroup A) : #\label{demo4:coe:ag:monoid:class}#
  Monoid A := Monoid.Class A (AbelianGroup.Monoid_of_Type_mixin A c).
Coercion AbelianGroup_to_Monoid (A : AbelianGroup.type) : Monoid.type := #\label{demo4:coe:ag:monoid:type}#
  Monoid.Pack (AbelianGroup.sort A) (AbelianGroup.class A).

Definition opp {A : AbelianGroup.type} : A -> A := ... .
(* Two Abelian group axioms `addrC` and `addNr` are omitted. *)
\end{coqcode}

Generally, a coercion from a structure X to another structure Y can (and should) have the following form, thanks to the corresponding coercion between classes \coq{X_class_to_Y_class}.
\begin{coqcode}
Coercion X_to_Y (T : X.type) : Y.type :=
  Y.Pack (X.sort T) ((* X_class_to_Y_class _ *) (X.class T)).
\end{coqcode}

@@

\begin{coqcode}
#\hbmixin{}# Record SemiRing_of_Monoid S
  of Monoid S := {
  one : S;
  mul : S -> S -> S;
  (* 7 axioms are omitted. *)
}.

#\hbfactory{}# Record Ring_of_AbelianGroup R
  of AbelianGroup R := {
  (* 2 operators and 5 axioms are omitted. *)
}.
#\hbbuilders{}# Context R (f : Ring_of_AbelianGroup R).
..
#\hbend{}#.
\end{coqcode}

Since semirings and Abelian groups do not inherit from each other,
the definition of semirings and Abelian groups are quite similar:
\begin{coqcode}
Module SemiRing.
Record axioms (S : Type) : Type := Class {
  Monoid_of_Type_mixin : Monoid_of_Type S;
  SemiRing_of_Monoid_mixin : SemiRing_of_Monoid S Monoid_of_Type_mixin;
}.
Structure type : Type := Pack { sort : Type; class : axioms sort; }.
End SemiRing.
\end{coqcode}
We define implicit coercions from the semiring structure to the carrier
and the monoid structure, and we lift semiring operators and axioms as follows:
\begin{coqcode}
Coercion SemiRing.sort : SemiRing.type >-> Sortclass.
Coercion SemiRing_to_Monoid : SemiRing.type >-> Monoid.type.

Definition one {S : SemiRing.type} : S := ... .
Definition mul {S : SemiRing.type} : S -> S -> S := ... .
(* 7 semiring axioms are omitted. *)
\end{coqcode}

The class record is defined by gathering the monoid, Abelian group, and
semiring \mixins{}. Since the rings inherit from monoids, semirings, and
Abelian groups, we define implicit coercions from the ring structure to those
three structures.
\begin{coqcode}
Module Ring.
Record axioms (R : Type) : Type := Class {
  Monoid_of_Type_mixin : Monoid_of_Type R;
  SemiRing_of_Monoid_mixin : SemiRing_of_Monoid R Monoid_of_Type_mixin;
  AbelianGroup_of_Monoid_mixin : AbelianGroup_of_Monoid R Monoid_of_Type_mixin; }.
Structure type : Type := Pack { sort : Type; class : axioms sort; }.
End Ring.

@@

Another example of elaboration problem is \(- 1\), which hides the term~\coq{opp _ (one _)}.
Here \coq{opp} and \coq{one} are respectively from Abelian groups and semirings,
which do not inherit each other but whose smallest common substructure is
rings; thus we have to extend the unifier to solve a unification problem
\(\coq{AbelianGroup.sort}~?_{\texttt{AbG}}~\coq{= SemiRing.sort}~?_{\texttt{SR}}\) by instantiating \(?_{\texttt{AbG}}\) and \(?_{\texttt{SR}}\) with \(\coq{Ring_to_AbelianGroup}~?_{\texttt{R}}\) and \(\coq{Ring_to_SemiRing}~?_{\texttt{R}}\) respectively where \(?_{\texttt{R}}\) is a fresh unification variable of type \coq{Ring.type}.
This hint can be given as follows:% can be done by the following canonical declaration.
\begin{coqcode}
Canonical AbelianGroup_to_SemiRing (S : Ring.type) :=
  SemiRing.Pack (AbelianGroup.sort (Ring_to_AbelianGroup S)) (Ring.class S)).
\end{coqcode}

@@

(* Bottom mixin in Fig. 2. *)
HB.mixin Record Monoid_of_Type M := {
  zero : M;
  add : M -> M -> M;
  addrA : associative add;
  add0r : left_id zero add;
  addr0 : right_id zero add;
}.
HB.structure Definition Monoid := { M of Monoid_of_Type M }.
Notation "0" := zero : hb_scope.
Infix "+" := (@add _) : hb_scope.

@@

(* Bottom left mixin in Fig. 2. *)
HB.mixin Record SemiRing_of_Monoid S of Monoid S := {
  one : S;
  mul : S -> S -> S;
  mulrA : associative mul;
  mul1r : left_id one mul;
  mulr1 : right_id one mul;
  mulrDl : left_distributive mul add;
  mulrDr : right_distributive mul add;
  mul0r : left_zero zero mul;
  mulr0 : right_zero zero mul;
}.
HB.structure Definition SemiRing := { S of Monoid S & SemiRing_of_Monoid S }.
Notation "1" := one : hb_scope.
Infix "*" := (@mul _) : hb_scope.

Lemma addrN {A : AbelianGroup.type} : right_inverse (zero : A) opp add.
Proof. by move=> x; rewrite addrC addNr. Qed.

(* Top right factory in Fig. 2. *)
HB.factory Record Ring_of_AbelianGroup R of AbelianGroup R := {
  one : R;
  mul : R -> R -> R;
  mulrA : associative mul;
  mul1r : left_id one mul;
  mulr1 : right_id one mul;
  mulrDl : left_distributive mul add;
  mulrDr : right_distributive mul add;
}.

@@

(* Top left factory in Fig. 2. *)
(* It is an exact copy of the bottom right mixin. *)
HB.factory Definition Ring_of_SemiRing R of SemiRing R := AbelianGroup_of_Monoid R.
(* The corresponding builder is the identity. *)
HB.builders Context (R : Type) (f : Ring_of_SemiRing R).
  Definition to_AbelianGroup_of_Monoid : AbelianGroup_of_Monoid R := f.
  HB.instance R to_AbelianGroup_of_Monoid.
HB.end.

(* Right-most factory in Fig. 2. *)
HB.factory Record Ring_of_Monoid R of Monoid R := {
  one : R;
  opp : R -> R;
  mul : R -> R -> R;
  addNr : left_inverse zero opp add;
  addrN : right_inverse zero opp add;
  mulrA : associative mul;
  mul1r : left_id one mul;
  mulr1 : right_id one mul;
  mulrDl : left_distributive mul add;
  mulrDr : right_distributive mul add;
}.
